---
title: "The rptPlus Sample Vignette"
author:
- name: Boris Steipe
  affiliation: 
  - http://orcid.org/0000-0002-1134-6758
  - University of Toronto
  email: boris.steipe@utoronto.ca
date: "`r Sys.Date()`"
abstract: >
  The rptPlus Sample Vignette demonstrates markdown syntax for the creation of a Bioconductor-compatible Vignette, and serves as a template that can be customized for your own contents.
output:
  BiocStyle::html_document:
    toc: true
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{The rptPlus Sample Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!--- this inline style is supports diff-like highlighting -->
<style>
  .add {
    color: #22863a;
    background-color: #f0fff4;
  }
  .del {
    color: #b31d28;
    background-color: #ffeef0;
  }
</style>


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
# knitr knits in a new session with empty global workspace. To make
# your package functions available in the vignette, you have to load
# the library. The following two lines should accomplish this without manual
# intervention:
pkgName <- trimws(gsub("^Package:", "", readLines("../DESCRIPTION")[1]))
library(pkgName, character.only = TRUE)

```


# About this Vignette

This sample Vignette explains how to customize this template for your own R package, and it contains hints on how to create meaningful contents. Thus the Vignette is a template for your own development - both in how it integrates with the rest of the package framework, and what you can do with it. Edit it to your needs. But do make sure your package **has** a Vignette! Obviously, the code and data assets contain comment headers throughout, and these are used by roxygen2 to compile package documentation and man pages. However, **documenting what the functions do does not necessarily explain how to use a package**. 

The `.Rmd` format in which this Vignette is written is a very clever way to enable ["literate programming"](https://en.wikipedia.org/wiki/Literate_programming): the source combines [markdown](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf) formatted text with executable R code, and the results of the code are integrated  into the resulting document. This magic is made possible by the fabulous [knitr package](https://en.wikipedia.org/wiki/Knitr), originally written by [Yihui Xie](https://yihui.name/en/about/) and integrated into the RStudio IDE. 

`knitr` can render the Rmarkdown contents of this Vignette to a variety of different formats, including html and pdf. Which format is targeted is defined in the metadata block at the top of the Vignette.  Specifically, the format used here is
["`BiocStyle::html_document`"](https://bioconductor.org/packages/devel/bioc/vignettes/BiocStyle/inst/doc/AuthoringRmdVignettes.html), which is a requirement for Bioconductor packages. An alternative format that is often used in CRAN and other packages is ["`html_vignette`"](https://bookdown.org/yihui/rmarkdown/r-package-vignette.html); this format aims to produce a vignette that is as small as possible.

&nbsp;

# Develop your Vignette

In RStudio, open the `./vignettes` folder and rename `rptPlus.Rmd` to `&lt;your-package-name&gt;.Rmd. Then open the file to edit it. Whenever you make changes, click on the **Knit** icon at the top of the editing pane, this will save the file, "knit" the output to an html file, and open the file in the viewer. You can also copy the path of the resulting file from the **R Markdown** tab in the console pane and open it in your normal browser.


## Customize the metadata

Edit the following information in the metadata block:

### Your Vignette Title

In the header block, change your Vignette title and the Vignette index entry:

<pre>

-   title: "<span class="del">The rptPlus Sample Vignette</span>"
+   title: "<span class="add">&lt;your-Vignette-title&gt;</span>"

</pre>

<pre>

vignette: >
  %\VignetteIndexEntry{<span class="del">The rptPlus Sample Vignette</span>}
  %\VignetteIndexEntry{<span class="add">&lt;your-Vignette-title&gt;</span>}
[...]

</pre>

### Author and Affiliation

The author name and affiliation in the document header are also generated from metadata. You can list several authors with their shared and unique affiliations and provide a linked email address to the corresponding author. For details, see [here](https://bioconductor.org/packages/release/bioc/vignettes/BiocStyle/inst/doc/AuthoringRmdVignettes.html#31_author_affiliations).

<pre>
author:
-   - name: <span class="del">Boris Steipe</span>
+   - name: <span class="add">&lt;your-name&gt;</span>
     affiliation: 
-     - http://orcid.org/<span class="del">0000-0002-1134-6758</span>
+     - http://orcid.org/<span class="add">&lt;your-ORCID-ID&gt;</span>
-     - name: <span class="del">University of Toronto</span>
+     - name: <span class="add">&lt;your-affiliation&gt;</span>
-     email: <span class="del">boris.steipe@utoronto.ca</span>
+     email: <span class="add">&lt;corresponding-author's-email&gt;</span>

</pre>


### Abstract and Running Headers

An abstract will be processed from the metadata block.

<pre>

abstract: >
-    <span class="del">The rptPlus Sample Vignette demonstrates[...]</span>
+    <span class="add">&lt;your-Vignette-summary&gt;</span>

</pre>


## Contents

### Text

Text in this Vignette is written in R Markdown format. 

<!-- quotation -->
> "Science is built up with facts, as a house is with stones. But a collection of facts is no more a science than a heap of stones is a house."
_(Henri Poincaré: La Science et l'hypothèse.)_

### Math

R markdown supports the rendering of equations with LaTeX markup. Here is an example:

The golden ratio $\phi$ <!-- this is an inline equation: single $ delimiters -->is 

$$ \frac{\left(1 + \sqrt{5}\right)}{2} $$ <!-- this is a block equation: double $$ delimiters -->

A "golden spiral" [^1] <!-- here is a footnote --> is a logarithmic spiral whose radius diminishes by $\frac{1}{\phi}$ every quarter turn. It is thus inscribed into a "golden rectangle". The golden spiral with initial radius 1 is defined by the polar equation
<!-- the preferred way to write block equations is to use latex commands directly, and label each equation block. BiocStyle then automatically adds equation numbers.-->

\begin{equation}
  r = \phi^{-\theta\frac{2}{\pi}}
  (\#eq:phiPolar)
\end{equation}


[^1]: <!-- Footnote text. This footnote is rendered as a margin note courtesy of BiocStyle -->cf. ["Golden spiral" on Wikipedia](https://en.wikipedia.org/wiki/Golden_spiral).

### Figures

The block of R code below is executed when the document is knit, and the resulting plot is  shown. The code first sets up `N` angles `theta` and then computes the radius according to  \@ref(eq:phiPolar) <!-- here is a back-reference to equation (1) -->. Polar coordinates are then converted to cartesian coordinates and scaled to the canonical golden rectangle. We plot an empty frame, draw the rectangle, draw segments for the enclosed golden sections, and finally draw the line of the spiral.

```{r, fig.cap="The golden spiral is shown in its enclosing golden rectangle."}
phi <- ((1 + sqrt(5)) / 2)^(1:-8)           # ten decreasing powers of phi
tStart <- -0.233
tEnd <- 8 * pi
N <- 800
theta <- seq(tStart, tEnd, length.out = N)
r = phi[1]^-(theta * (2  / pi))
x <- r * cos(theta)
y <- r * sin(theta)

x <- ((x + abs(min(x))) * phi[1]) / (max(x) - min(x))  # scale to [0, phi]
y <-  (y + abs(min(y)))           / (max(y) - min(y))  # scale to [0, 1]

oPar <- par(mar = rep(0.25, 4))             # set thin margins
plot(x, y,                                  # plot empty frame ...
     type = "n",
     xlim = c(0, phi[1]),
     ylim = c(0, 1),
     xlab = "", ylab = "", 
     axes = FALSE,
     asp = 1)                               # with fixed aspect ratio

rect(0, 0, phi[1], 1, border = "#CCCCCC")   # draw enclosing rectangle


C <- "#C9C9EE"                              # draw enclosed golden sections
segments(phi[1]-phi[2],            0,phi[1]-phi[2],       phi[2],col=C,lwd=0.4)
segments(            0,phi[2]-phi[3],phi[1]-phi[2],phi[2]-phi[3],col=C,lwd=0.4)
segments(phi[2]-phi[3],            0,phi[2]-phi[3],phi[2]-phi[3],col=C,lwd=0.4)
segments(phi[2]-phi[3],phi[3]-phi[4],phi[1]-phi[2],phi[3]-phi[4],col=C,lwd=0.4)
segments(phi[4]+phi[7],phi[3]-phi[4],phi[4]+phi[7],phi[2]-phi[3],col=C,lwd=0.4)
segments(phi[2]-phi[3],phi[5]+phi[8],phi[4]+phi[7],phi[5]+phi[8],col=C,lwd=0.4)
segments(phi[4]+phi[8],phi[3]-phi[4],phi[4]+phi[8],phi[5]+phi[8],col=C,lwd=0.4)
segments(phi[4]+phi[8],phi[5]+phi[9],phi[4]+phi[7],phi[5]+phi[9],col=C,lwd=0.4)

lines(x, y, col = "#CC0000")               # draw spiral

par <- oPar                                # reset graphics state

```


### Tables

Printing tables is straightforward using the defaults in `knitr::kable()`. But getting the data from your package can be tricky: the package needs to be built and installed first and loaded in the setup chunk of the Vignette. Then you can simply refer to exported data in the package namespace. However the usual `system.file()` will still fail. Not to worry: the correct path is simply `../inst/extdata/<fileName>` since knitr knits in `./vignettes` as its working directory. 

Example 1. The standard genetic code is provided as sample data with `rptPlus`:

```{r, echo=FALSE, results='asis'}
x <- matrix(paste0(names(rptGC), ": ", rptGC), ncol=4)
nuc <- c("A", "G", "C", "T")
colnames(x) <- paste0(nuc, "..", sep ="")
rownames(x) <- paste0(".", rep(nuc, each = 4), nuc, sep ="")

knitr::kable(x,
             caption = "The standard genetic code.",
             align = "c")
```

Example 2. Printing the values in `./inst/extdata/test_lseq.dat`

```{r, echo=FALSE, results='asis'}
x <- readLines("../inst/extdata/test_lseq.dat")
knitr::kable(x, caption = "Five log-spaced numbers in [1, 10]")
```


# Further reading

- The [Rstudio Markdown pages](https://rmarkdown.rstudio.com/)
- The [RStudio R markdown cheat sheet](https://www.rstudio.com/wp-content/uploads/2016/03/rmarkdown-cheatsheet-2.0.pdf)
- The [RStudio R markdown reference guide](https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf)
- [Detailed **"chunk**" options](https://yihui.name/knitr/options/)
- The [Bioconductor LaTEX style](https://bioconductor.org/packages/release/bioc/vignettes/BiocStyle/inst/doc/LatexStyle2.pdf)

<!-- End -->
